---
title: "hw6_p8105_jld2227"
author: "James Dalgleish"
date: "November 20, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(magrittr)
library(modelr)
```

Instruction: Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake.Modifiy victim_race to have categories white and non-white, with white as the reference category. Be sure that  victim_age is numeric.

In this dataset, we'll look at the odds ratios of homicide victims by city and deduce, which cities have the highest risk of homicide for non-white victims as compared with white victims, with confidence intervals to get a sense of the plausible values of the odds ratio.

To begin with, we'll pull the homicide data directly from the washington post
github site, remove the odd looking characters from the data with iconv,
convert age to numeric, change the victim race to a white-nonwhite only factor, and
change this factor to have white as the reference category. We'll also create a city_state variable by combining city and state, identify which cases are unresolved (not closed by arrest). For a bit of a sanity check (to make sure numbers add up), we'll also add a variable for resolved cases. As such, the total cases should make up the resolved and unresolved cases, and they do. We'll also remove the non-existent city, Tulsa, AL, and Dallas, Phoenix, and Kansas City.
```{r data_import, message = F, warning = F}
homicide_data <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% 
  mutate( victim_first = iconv(x = victim_first, from = "latin1", to =  "ASCII",sub = ""),            victim_last = iconv(x = victim_last, from = "latin1", to =  "ASCII",sub = ""),
          victim_age = as.numeric(victim_age), 
          victim_race = ifelse(victim_race == "White","white","non-white") %>% 
          as.factor() %>% 
          forcats::fct_relevel("white","non_white"), 
          city_state = str_c(city,", ",state), 
          unresolved = disposition %in% c("Closed without arrest","Open/No arrest"),
          resolved = disposition %in% c("Closed by arrest")) %>% 
  filter(!(city_state %in% 
  c("Dallas, TX","Phoenix, AZ","Kansas City, MO","Tulsa, AL")))

homicide_data %>% 
  head() %>% 
  knitr::kable()
```



Instruction: For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

Now, we'll focus on Baltimore. We'll do this by filtering for cases only from baltimore, fitting a logistic regression model for resolved cases with age, sex, and race of the victim as predictors, then create confidence intervals using broom::tidy and odds ratios by calculating $OR=e^{estimate}$ and doing the same to the lower and upper bound of that same estimate.

```{r baltimore, warning = F}
balt_hom_fit <- homicide_data %>%
   filter(city_state == "Baltimore, MD") %$%
  glm(resolved ~ victim_age + victim_sex + victim_race,
      family = binomial())
balt_city_ci_or <- balt_hom_fit %>%
  broom::tidy(conf.int=TRUE) %>%
  mutate(OR = exp(estimate), OR_conf.low = exp(conf.low), OR_conf.high = exp(conf.high))
balt_city_ci_or %>% 
  knitr::kable()
```
Instruction: Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

Now we'll grab the odds ratios for resolved white to nonwhite for all cities, not just baltimore.
We'll do this by grouping by each City, State combination, nesting the individual City, State groups, then running a general linear model on the nested data (a column of tibbles),
then we'll tidy the model output and produce confidence intervals.  We'll remove the nested data and unused glm objects from the tibble, unnest.  Utilizing the estimates, we'll calculate odds ratios again, for each city, under the same model. We'll filter these estimates for only those for the race factor (white-nonwhite) and following this, we'll display the top odds ratio city_states.
```{r all_cities, warning = FALSE}
all_city_glm <- homicide_data %>%
  group_by(city_state)  %>%
  nest() %>% 
  mutate(hom_glm = map(data,~glm(formula = resolved ~ victim_age + victim_sex + victim_race,data=.x,family = binomial()))) %>% 
         mutate(hom_tidy = map(hom_glm,broom::tidy,conf.int=TRUE))  %>% 
  select(-data,-hom_glm) %>% 
  unnest() %>% 
  mutate(OR = exp(estimate), OR_conf.low = exp(conf.low), OR_conf.high = exp(conf.high) ) %>% 
  filter(term == "victim_racenon-white")
all_city_glm %>% 
  arrange(-OR) %>% 
  head() %>% 
 knitr::kable()  
```

Instruction: Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

Now, we'll visualze the ORs and CIs, sorted by the odds ratio value along the x axis.  We'll first accomplish this by changing the order of the odds ratio factor, then plotting the odds ratio along with it's confidence interval.
```{r plot}
all_city_glm %>% 
  mutate(city_state = forcats::fct_reorder(city_state,OR)) %>% 
  ggplot(data = .,
  aes(x = city_state, y = OR, ymin = OR_conf.low, ymax = OR_conf.high)) +
  geom_point() +
  geom_errorbar() +
  theme(axis.text.x = element_text(angle = 270)) +
  labs(x = "City, State",
       title = "Odds Ratio Estimate, White-Nonwhite for unresolved cases",
       y = "Odds Ratio & Confidence Interval")
```
Instruction: Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

To create a predictive model of birthweight, which is an indicator of newborn health, we'll examine a small dataset of children with several variables recorded for them, including maternal variables such as the mother's age at menarche, female race, number of live births, as well as paternal variables. The dataset also contains physical measurements of the baby, which likely have the strongest impact on the birthweight.

We'll now read in the birthweight data, being careful to change the sex to a factor variable with "male" and "female", and we'll also make factors out of father's race, mother's race, prescnce of malformations in the child. The rest of the variables are in their appropriate form. We also note there's no missing data from the table generated below.
```{r}
birthweight <- read_csv("http://p8105.com/data/birthweight.csv", col_types = "cnnnnncnlnnncnnnnnnn") %>%
  mutate(babysex = as.factor(babysex)) %>% 
  mutate(babysex = recode(babysex, `1` = 'male', `2` = 'female'))  %>% 
  janitor::clean_names() %>% 
  mutate(frace = as.factor(frace)) %>% 
  mutate(malform = as.factor(malform)) %>% 
  mutate(mrace = as.factor(mrace))
  skimr::skim(birthweight)
```

Instruction: Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two.  Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Perhaps the best way to build a model is with some user input based on hypothesis to begin with. Factors that do not affect birthweight significantly or have small effect sizes can be removed. We should do this with some sense of the distribution of the data (whether some variables are skewed in distribution or normal) and the linear relataionships between variables in the dataset to each other (correlations). 

Our initial, hypothesis based model will include physical attributes of the newborn, including the length of the infant and the circumference of the head, as well as the age of the child at birth (in gestational weeks) as full-term infants are larger than premature ones. A priori, we would think that the number of low birth weight babies that each mother had would be helpful, but as we'll see, no one sampled had any low weight children prior to this study and it's a useless variable in this particular dataset.

Here's our initial model, including the terms previously stated. We'll note that the number of low weight children for each mother of the individual children in the study simply drops out of the model.
```{r initial_model}
initreg <- birthweight %>%
  glm(data = ., bwt ~ bhead + blength + gaweeks + pnumlbw)
initreg %>% 
  broom::glance()

```
Why it dropped out of the model becomes clear in the summaries below. pnumlbw ("previous number of low birth weight babies") and pnumsga ("number of prior small for gestational age babies") are zero from minimum to maximum. Both of these variables should not be included in the model.
```{r}
skimr::skim_with(numeric = list(hist = NULL, min = min, max = max, kurtosis = e1071::kurtosis, skewness = e1071::skewness, IQR = IQR,
mean = mean,
cv = function(x){sd(x)/mean(x)})) 
skimr::skim(birthweight) %>% 
  skimr::kable() 
```
We also note that there is some mild skewness, but only extreme in the parity variable ("number of live births prior to this pregnancy"). To do a little exploratory data analysis, we'll look at the correlations among all variables using a colored corrplot. The deleted observations have no correlation values as they are entirely columns composed of zeros.
```{r}
  dplyr::select_if(.tbl = birthweight,.predicate = is.numeric)  %>% 
  cor() %>% 
corrplot::corrplot()
```
From this, we might notice that ppbmi("mother’s pre-pregnancy BMI"), ppwt("mother’s pre-pregnancy weight (pounds)"), and delwt("mother’s weight at delivery (pounds)") are all correlated. This suggests that delwt may be less important in prediction than other variables as the delivery weight of the mother may have a lot to do with her personal weight and BMI, rather than the child she delivers. Despite this, as it turns out that delivery weight is linearly related to birth weight, although not as strongly as other variables.

To focus on variables that are linearly related to birthweight, we could also produce a list of variables and their sorted correlations with the predictor. Doing so reveals that delwt (delivery weight) and maternal weight gain (wtgain) might be added to the model as they have some of the strongest relationships with the predictor, despite having relationships with the maternal weight and bmi.

To do this, we have to select only the columns that are numeric, create a correlation tibble with the name of the correlated variables and just select the name of the variable and it's correlation with birthweight, then sort by birthweight with the largest correlations at the top.
```{r}
dplyr::select_if(.tbl = birthweight,.predicate = is.numeric)  %>% 
  cor() %>% 
  as.data.frame() %>% 
    add_rownames() %>% 
  as.tibble() %>% 
  select(rowname,bwt) %>% 
  arrange(-bwt)
```
Knowing this, we'll add delivery weight and maternal weight gain to the model and see our AIC decrease a bit, suggesting an improved  model.
```{r}
initreg_plus_2 <- birthweight %>%
  glm(data = ., bwt ~ bhead + blength + gaweeks + delwt + wtgain) 
initreg_plus_2 %>% 
  broom::glance()

```
As a final step, we'll examine a full additive model and see if there are any variables that might be included as well, based on their effect sizes and p-values.
```{r}
birthweight %>%
  glm(data = ., bwt ~ .) %>% 
  broom::tidy() %>% 
  arrange(p.value)
```
'Smoken' is the average number of cigarettes per day and seems to have a definite linear relationship with the outcome variable. What's further, these findings could have important health implications given the linear relationship is fairly strong.  We'll add this and compare all models discussed, along with the full and intercept only models for comparison. 
```{r}
smoke_model <- birthweight %>%
  glm(data = ., bwt ~ bhead + blength + gaweeks + delwt + wtgain + smoken)
full_model <- birthweight %>%
  glm(data = ., bwt ~ .) 
int_model <- birthweight %>%
  glm(data = ., bwt ~ 0) 

```
To display the model improvement through this iterative process, we'll plot the AIC for each model. We'll do this by turning the models into a list of models, performing a summary with broom::glance that will get the AIC for each model and put it into a dataframe with map_dfr, then create a variable with the labels for the plot along the x axis, clean the names of the variables, and put the model name at the left hand side of the dataframe.
```{r aic_plot}
model_tibble <- list(int_model,initreg,initreg_plus_2,smoke_model,full_model) %>% 
  map_dfr(broom::glance) %>% 
  mutate(model_name = forcats::fct_inorder(c("intercept_only","head+len+wks","+delwt+wtgain","+smoke","full"))) %>%  
  janitor::clean_names() %>% 
  select(model_name,everything())
model_tibble %>% 
  knitr::kable()
```
As we'll notice, all the models have an AIC that is MUCH lower than the intercept, suggesting our variables are indeed better than the model with no variables other than the intercept. We'll plot every model except for the intercept below, finding that although the model we ended with is great, it is not as optimal as the model including all variables. We'll color the points by the logliklihood of the model as well using thte viridis colorscale, produce points and lines to show the trend, and label the points using ggrepel, which stops points from overlapping with text.
```{r}
library(ggrepel)
model_tibble %>%
  filter(model_name != "intercept_only") %>% 
  ggplot(aes(x=model_name, y = aic, group = 1)) + 
  geom_point(aes(fill=log_lik)) +
  geom_line(aes(x = model_name, y =  aic)) + 
  geom_text_repel(aes(label = aic)) +
  labs(x = "Model", y = "AIC value, log scale", title = "Birthweight Models") +
  scale_fill_viridis_c()
```
Instruction: Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.
```{r}
birthweight %>% 
lm(data = ., bwt ~ bhead + blength + gaweeks + delwt + wtgain + smoken) %>% 
  modelr::add_predictions(data = birthweight, model=.) %>% 
  modelr::add_residuals(data = ., model=smoke_model) %>% 
  ggplot(aes(x = pred,y = resid)) +
  geom_point() +
  labs(x = "model residuals",
       y = "fitted (predicted) values",
       title = "Residuals vs Fitted",
       subtitle = "bwt ~ bhead + blength + gaweeks + delwt + wtgain + smoken")
```
From the plot, we can see that there is some slight fanning in the low residual values due to some outliers in the data, but otherwise, the residuals seem to be evenly distributed about zero.
Instruction: Compare your model to two others:
One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Instruction: Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.
```{r cross_validation}
length_age_mod <- birthweight %>% 
  glm(data = ., bwt ~ blength + gaweeks )
head_len_sex_mod <- birthweight %>% 
  glm(data = ., bwt ~ bhead*blength*babysex )
set.seed(1)
birthweight %>% 
  modelr::crossv_mc(100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble)) %>% 
  mutate(length_age_mod = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         head_len_sex_mod = map(train, ~lm(bwt ~ bhead*blength*babysex, data = .x)),
         smoke_mod = map(train, ~lm(bwt ~ bhead + blength + gaweeks + delwt + wtgain + smoken, data = .x))) %>% 
  mutate(rmse_la    = map2_dbl(length_age_mod, test, ~rmse(model = .x, data = .y)),
         rmse_hls = map2_dbl(head_len_sex_mod, test, ~rmse(model = .x, data = .y)),
         rmse_smoke = map2_dbl(smoke_mod, test, ~rmse(model = .x, data = .y))) %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  ggplot(aes(x = model, y = rmse, color = model)) +
geom_violin() +
geom_point() +
  labs(x = "model",
       y = "Root Mean Squared Error", 
       title = "Distribution of RMSE values in 100 fold Cross Validation, by model")
```

