---
title: "hw6_p8105_jld2227"
author: "James Dalgleish"
date: "November 20, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
```

instruction: Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race.
Also omit Tulsa, AL – this is a data entry mistake.
Modifiy victim_race to have categories white and non-white, with white as the reference category.
Be sure that  victim_age is numeric.
```{r data_import}

homicide_data <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% #read in data.
  mutate( victim_first = iconv(x = victim_first, from = "latin1", to =  "ASCII",sub = ""), #fixes bad characters in victim name
          victim_last = iconv(x = victim_last, from = "latin1", to =  "ASCII",sub = ""),
          victim_age = as.numeric(victim_age), #converts victim age to numeric.
          victim_race = ifelse(victim_race == "White","white","non-white") %>% #binarizes the categories to white or nonwhite, depending on if the race is "White" or not.
            as.factor() %>% #converts to factor
            forcats::fct_relevel("white","non_white"), #sets reference cat to white
          #This removes nonstandard characters that cause problems with skimr.
          city_state = str_c(city,", ",state), #create city state variable by string concatenation.
          unresolved = disposition %in% c("Closed without arrest","Open/No arrest"),
          #creates a variable for whether the case was  unresolved.
          resolved = disposition %in% c("Closed by arrest")) %>% 
#creates a variable for whether the case was resolved, to double check the numbers.
#Included the resolved cases as a separate variable to clearly show that the resolved and unresolved cases sum to the total (and hence the calculations make sense).
  filter(!(city_state %in% c("Dallas, TX","Phoenix, AZ","Kansas City, MO","Tulsa, AL")))

```



For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object;
apply the broom::tidy to this object;
and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r}
hom_fit <- homicide_data %>%
   filter(city_state == "Baltimore, MD") %$%
  glm(resolved ~ victim_age + victim_sex + victim_race,
      family = binomial())
balt_city_glm <- hom_fit %>%
  broom::tidy(conf.int=TRUE) %>%
  mutate(OR = exp(estimate))
all_city_glm <- homicide_data %>%
  group_by(city_state)  %>%
  nest() %>% 
  mutate(hom_glm = map(data,~glm(formula = resolved ~ victim_age + victim_sex + victim_race,data=.x,family = binomial()))) %>% 
         mutate(hom_glm = map(hom_glm,broom::tidy,conf_int=TRUE))  %>% 
  select(-data) %>% 
  unnest() %>% 
  mutate(OR = exp(estimate))
all_city_glm %>% 
 knitr::kable()  
```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r plot}
all_city_glm %>% 
  ggplot(data= .,
         aes(x = city_state, y = estimate, ymin = conf.low, ymax = conf.high))
  geom_point() +
    geom_errorbar()
    
  #confidence intervals with dots.

```

